---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Number 1: Loading and processing the data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(ggplot2)
```

### Load the data

```{r}

originalData <- read.csv("~/R/R Course/Course 5/repdata_data_activity/activity.csv")
```

### Remove NAs from the data

```{r}

activityData <- originalData[complete.cases(originalData), ]
```


## Number 2: What is mean total number of steps taken per day?

### Group data by day
```{r}

byday <- group_by(activityData, date)

stepsPerDay <- summarize(byday, total = sum(steps, na.rm = TRUE))

```

### Create a histogram showing total number of steps per day

```{r}

hist(stepsPerDay$total, main = "Total Steps Taken per Day", xlab = "Number of Steps per Day", ylab = "Frequency")

```

### Calculate mean and median of steps per day

```{r}

mean(stepsPerDay$total)
median(stepsPerDay$total)

```

## Number 3: What is the average daily activity pattern?

### Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis).

#### First, calculate average steps per interval

```{r}

avgStepsPerInterval <- aggregate(steps ~ interval, activityData, mean)
```

#### Second, plot data as a time series plot

```{r}

plot(avgStepsPerInterval$interval, avgStepsPerInterval$steps, type = "l", main = "Average Steps Taken Across All Days", xlab = "5 Minute Interval", ylab = "Average Number of Steps Taken")

```

### Find which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps

#### First, find maximum steps interval index

```{r}

maximumSteps <- which.max(avgStepsPerInterval)

```

#### Second, find values for maximum steps interval index

```{r}

avgStepsPerInterval[maximumSteps,]

```


## Number 3: Imputing missing values

### Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}

allNAs <- sum(is.na(originalData$steps))
allNAs

```

### Create a function to return mean based on 5-minute interval


```{r}
getavgPerInterval <-function(interval){
    avgStepsPerInterval[avgStepsPerInterval$interval==interval,]$steps
}

```

### Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
newActivityData <- originalData

for (i in 1:nrow(newActivityData)) {
    if(is.na(newActivityData[i,]$steps)) {
        newActivityData[i,]$steps <- getavgPerInterval(newActivityData[i,]$interval)
    }
}
```

### Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day

```{r}
newStepsPerDay <- aggregate(steps ~ date, newActivityData, sum)

hist(newStepsPerDay$steps, main = "Total Steps Taken per Day (NAs replaced)", xlab = "Number of Steps per Day", ylab = "Frequency")

```


## Number 4: Are there differences in activity patterns between weekdays and weekends?

### Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day

```{r}

newActivityData$date <- as.Date(strptime(newActivityData$date, format="%Y-%m-%d"))
newActivityData$day <- weekdays(newActivityData$date)
```

#### Create a fuunction to differentiate weekday and weekend - "If the day of week (DOW) is Saturday or Sunday mark as weekend, Else mark as weekday"

```{r}

for (i in 1:nrow(newActivityData)) {
  if(newActivityData[i,]$day %in% c("Saturday","Sunday")) {
    newActivityData[i,]$day <- "weekend"
  }
  else{newActivityData[i,]$day <- "weekday"
  }
}

```

#### Aggregate the average steps per day of week

```{r}
avgStepsPerDOW <- aggregate(steps ~ interval + day, newActivityData, mean)
```

### Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r}
ggplot(avgStepsPerDOW, aes(x=interval, y=steps, color =day)) + geom_line() + labs(title="Avg Steps Taken Across All Weekday or Weekend days", x= "Interval", y="Number of Steps") + facet_wrap(~day, ncol = 1, nrow =2)

```
